Reintroduction of resistant frogs facilitates landscape-scale recovery in the presence of a lethal fungal disease

### 2024-04-09 doc/manuscript/translocation.qmd
Update sensitivity text in ms

Results from the individual variant and splined window analyses show that recovering populations have signatures of selection on multiple regions of the genome.
The analysis of individual variants identified 11 "outlier" SNPs (i.e., showing significantly different allele frequencies between naive versus recovering populations) from 7 distinct genes across 4 contigs (@fig-selectionresults B, C SI).
One of the 7 identified genes (LOC108802036) does not have an associated annotation.
For the outlier SNPs, frequency differences between the naive and recovering populations ranged from 0.41 to 0.86.
Most of these SNPs showed frequency differences in only a subset of the sampled populations (@fig-allelefrequencies A, B), but the SNP in the RIN3 gene showed consistent differences in frequencies across all populations (@fig-allelefrequencies C).
This is suggestive of parallel selection at this locus across multiple populations.
The other 6 outlier variants showed less consistent frequency differences across the study populations, but for these we still found a statistically significant signal of selection in 2 of the 3 genetic clusters (containing populations 5–9; @fig-allelefrequencies, @fig-selectionresults A SI).
Therefore, although some outlier variant associations have a more limited geographic extent than RIN3, they still describe results that suggest parallel evolutionary changes following Bd exposure.

@@

After parameterizing our model with CMR-estimated adult frog survival probabilities and other known vital rates ([@tbl-param_values]), we performed four analyses.
First, we compute the long-run growth rate $\lambda$ for each of our 12 translocated populations to determine if the populations were deterministically predicted to grow or decline in the long-run.
Second, we compute the elasticity of $\lambda$ to four key model parameters to quantify how much changes in these parameters affected the long-run growth rate (@fig-viability-supp SI).
This also helped us determine where in the model environmental variation in juvenile survival and recruitment would have the largest effects on population dynamics.
Third, we included demographic stochasticity and environmental stochasticity in $\sigma_{J_1}$ in our model and simulated the 50-year viability (i.e., 1 - extinction probability) of populations given an introduction of 40 adult individuals into an unoccupied habitat.
Finally, we fit our model to our longest translocation trajectory to confirm that our model could reasonably reproduce the observed recovery trajectories of MYL frogs following reintroductions.
Additional details are provided in **Supporting Information - Population viability modeling - Model analysis and simulation**.
Code to replicate the analyses can be found at <https://github.com/SNARL1/translocation>.

@@

Second, we performed a local elasticity analysis on $\lambda$ with respect to parameters $\sigma_{J_1}$, $\sigma_{J_2}$, $\sigma_{A_R}$, and $F$ to determine how small changes in these parameters could influence the long-run deterministic growth rate of populations (@fig-viability-supp SI).

@@

![Sensitivity analysis of the stage-structured MYL frog model. Elasticity of $\lambda$ with changes in six parameters: $\sigma_{A_R}$ (yearly survival probability of naturally recruited adults), $F$ (number of eggs produced by a female frog in a year that successfully hatch), $\sigma_{J_1}$ (yearly probability of survival of year-1 juveniles that also affects recruitment), $\sigma_{J_2}$ (yearly probability of survival and recruitment of year-2 juveniles), $p_{L_2}$ (probability that year 2 tadpoles metamorph into juveniles), $p_{J_1}$ (probability that year 1 juveniles mature to adults). We did not include $p_{L_1}$ as this parameter is generally 1 across populations (Table S1) -- tadpoles must spend at least a year as tadpoles before metamorphosing. Elasticity is calculated at the default parameter values for each population and $\sigma_{J_1}$ = 0.09.](figures/pop_viability_figures_for_supp.jpg){#fig-viability-supp width="550"}

>>>>

Results from the individual variant and splined window analyses show that recovering populations have signatures of selection on multiple regions of the genome.
The analysis of individual variants identified 11 "outlier" SNPs (i.e., showing significantly different allele frequencies between naive versus recovering populations) from 7 distinct genes across 4 contigs (@fig-selectionresults B, C SI).
One of the 7 identified genes (LOC108802036) does not have an associated annotation.
For the outlier SNPs, frequency differences between the naive and recovering populations ranged from 0.41 to 0.86.
Most of these SNPs showed frequency differences in only a subset of the sampled populations (@fig-allelefrequencies A, B), but the SNP in the RIN3 gene showed consistent differences in frequencies across all populations (@fig-allelefrequencies C).
This is suggestive of parallel selection at this locus across multiple populations.
The other 6 outlier variants showed less consistent frequency differences across the study populations, but for these we still found a statistically significant signal of selection in 2 of the 3 genetic clusters (containing populations 5--9; @fig-allelefrequencies, @fig-selectionresults A SI).
Therefore, although some outlier variant associations have a more limited geographic extent than RIN3, they still describe results that suggest parallel evolutionary changes following Bd exposure.

@@

After parameterizing our model with CMR-estimated adult frog survival probabilities and other known vital rates ([@tbl-param_values]), we performed four analyses.
First, we compute the long-run growth rate $\lambda$ for each of our 12 translocated populations to determine if the populations were deterministically predicted to grow or decline in the long-run.
Second, we compute the elasticity and sensitivity of $\lambda$ to nine model parameters to quantify how much changes in these parameters affected the long-run growth rate (@fig-viability-supp SI).
This also helped us determine where in the model environmental variation in juvenile survival and recruitment would have the largest effects on population dynamics.
Third, we included demographic stochasticity and environmental stochasticity in $\sigma_{J_1}$ in our model and simulated the 50-year viability (i.e., 1 - extinction probability) of populations given an introduction of 40 adult individuals into an unoccupied habitat.
Finally, we fit our model to our longest translocation trajectory to confirm that our model could reasonably reproduce the observed recovery trajectories of MYL frogs following reintroductions.
Additional details are provided in **Supporting Information - Population viability modeling - Model analysis and simulation**.
Code to replicate the analyses can be found at <https://github.com/SNARL1/translocation>.

@@

Second, we performed a local elasticity and sensitivity analysis on $\lambda$ with respect to parameters $\sigma_{L_1}$, $\sigma_{L_2}$, $\sigma_{L_3}$, $\sigma_{J_1}$, $\sigma_{J_2}$, $\sigma_{A_R}$, $F$, $p_{L_2}$, $p_{J_1}$ to determine how small changes in these parameters could influence the long-run deterministic growth rate of populations (@fig-viability-supp SI).
Note that $\lambda$ is equally or more sensitive to changes in adult survival ($\sigma_{A_R}$) and survival of year-1 juveniles ($\sigma_{J_1}$) than it is to other model parameters, providing additional justification for focusing on variation in $\sigma_{A_R}$ and $\sigma_{J_1}$ in our viability analyses.

@@

![Sensitivity analysis of the stage-structured MYL frog model. A. Elasticity of $\lambda$ with changes in nine parameters: $\sigma_{A_R}$ (yearly survival probability of naturally recruited adults), $F$ (number of eggs produced by a female frog in a year that successfully hatch), $\sigma_{J_1}$ (yearly probability of survival of year-1 juveniles that also affects recruitment), $\sigma_{J_2}$ (yearly probability of survival and recruitment of year-2 juveniles), $p_{L_2}$ (probability that year-2 tadpoles metamorph into juveniles), $p_{J_1}$ (probability that year-1 juveniles mature to adults), $\sigma_{L_1}$ (survival probability of year-1 tadpoles), $\sigma_{L_2}$ (survival probability of year-2 tadpoles), and $\sigma_{L_3}$ (survival probability of year-3 tadpoles). B. Sensitivity of $\lambda$ with changes in the same nine parameters. We did not include $p_{L_1}$ as this parameter is generally 1 across the populations in this study (Table S1) -- tadpoles spend at least a year as tadpoles before metamorphosing. Elasticity and sensitivity is calculated at the default parameter values for each population and $\sigma_{J_1}$ = 0.09.](figures/pop_viability_figures_for_supp.jpg){#fig-viability-supp width="550"}
